Supplementary Information: Drought sensitivity is not a species trait: Tropical tree drought sensitivity is jointly shaped by drought characteristics, species adaptations, and individual microenvironments

### 2025-09-04 doc/response_to_reviewer_R1.Rmd

---
title: 'Responses to the Reviewer'
output: 
    bookdown::word_document2: 
    #   reference_docx: word-styles-reference.docx #make sure this line is indented! see this guide for formatting: https://rmarkdown.rstudio.com/articles_docx.html
      reference_docx: word-styles-reference-singlespace.docx #make sure this line is indented! see this guide for formatting: https://rmarkdown.rstudio.com/articles_docx.html
    #pdf_document:
    # engine: xelatex 
  #officedown::rdocx_document
bibliography: growth-precip-thailand.bib
# csl: new-phytologist.csl
csl: ecology-letters.csl 
---

**Title:** Drought response is not a species trait: tropical tree drought sensitivity is shaped by drought characteristics, species adaptations, and individual microenvironments

@@

Figure 1a: “Dy” is not correct

**RESPONSE: Thanks for noticing this error. We have fixed this to say "Day of year".**

Fig. 2. It is unnecessary to include the years 2023 and 2024 on the y-axis, as they were not evaluated. I think this figure could be analyzed further. For example, it took two years after 2014 for growth to recover to previous year (for ALL species) unlike in 2010 and 2020, when it rebounded within a year to levels equal to or higher than the previous year.

*RESPONSE: Thanks for this suggestion. We have removed 2023 and 2024 from the x-axis. While we agree that there is a lot more to discuss and analyse from this figure, we wanted to be careful not to overinterpret the growth patterns without adequate analysis of other covariates. For example, the pattern of suppressed growth after the 2015 drought is not necessarily because of lag times in recovery but more likely because drought conditions were prolonged into 2016 - this was a multi-year drought - and El Nino conditions switched to La Nina only late in 2016, leading to a late monsoon. Our analysis was focussed on episodic droughts, and therefore we chose not to analyse and interpret the whole timeseries in this study. We recognise however, that lagged responses and compensatory responses would be important in such timeseries and hope to answer these as the dataset accumulates.*

@@

Fig. S1 What does the shade area indicate? The three different drought events?

**RESPONSE: Yes, the shaded area indicates the three different drought years. We have edited the text to reflect this.**

Fig S6 The R Square is too low, so it is difficult to detect the linear relationship. The same issue for Fig. S11 and Fig. S12. For the non-significant correlation, it’s better use dashed lines instead of solid line.

**RESPONSE: Thanks for this suggestion. The dotted line in Fig S6 was originally the 1:1 line - a strict null expectation. We have now added dashed lines for the significant correlations. Following the same logic, for S11 and S12, we have now removed lines for the non-significant correlations.**

### 2025-09-08 doc/response_to_reviewer_R1.Rmd

•	Since the study is based on only one site, it is relevant to improve the discussion section with other already published study sites (add more references), mainly link the global implications, such mortality rate and vulnerability in tropical forests. See https://urldefense.proofpoint.com/v2/url?u=https-3A__doi.org_10.1126_science.adq6607&d=DwIFaQ&c=009klHSCxuh5AI1vNQzSO0KGjl4nbi2Q0M1QLJX9BeE&r=6fH468T58at0TtM59sMf109A_dRgo6XTfGMOVX5Ewrw&m=fPE_4y1UPsTl9St_Y44IUO7VIaQJOJIUiHtK15-Nr7g4UiVKbtYQ-TE0043SLAj6&s=SxqVZgFkbqAZ0-VEvdQntXWpGEUu9lr25rgxQtwaVdw&e=

*RESPONSE:*

•	There was something missing to discuss about the species in particular, some that have been more affected than others and why, especially those in conservation categories.

**RESPONSE: Thanks for this suggestion. We have now included a supplementary table that details the response of specific species to specific drought events. This also highlights our main message that the species responses are dissimilar acorss drought events.**

@@

2 The topographic wetness index. Are there any measured soil moisture data in this plot? Since this index is only based on topography, it will be good if the author could provide some measured soil moisture data to calibrate the prediction based on model prediction.

**RESPONSE: Unfortunately soil moisture data from the plot is not available for the years of study across which dendroband data are available. HKK, being fairly remote, has limited access to power sources and other infrastructure that are needed to maintain such data collection. We are currently piloting a few basic soil moisture sensors, however, this will have limited value for calibration as they only measure the top layer of the soil and are limited in spatial scale compared to the spatial resolution of TWI data.**

3 In the case of deciduousness, the authors indicated that they used the deciduous index from Williams et al. (2008). Since different species might shed leaves in different season. Did the authors consider the impact of the timing of deciduousness on the growth?

*RESPONSE: Thanks for this interesting comment. We agree that there are several hydraulic strategies both above and belowground that might impact tree growth sensitivity to drought. We chose deciduousness as a simple, illustrative example of a strategy that might be associated with some of this variation. We agree that the timing of deciduousness may be an important factor to consider, as might leaf turgor loss point, wood density etc. and considering these would be a key direction for future analyses, but beyond the scope of the current analysis.*

### 2025-09-08 doc/supplementary_information.Rmd
update with table for species sensitivity medians and IUCN status

>>>>

**Table S1: Species characteristics and their median sensitivities** for the three drought years. Species characteristics reported include accepted scientific names of species analysed, their ForestGEO codes, deciduousness values (amount of leaf loss at maximum) and IUCN status as accessed in September 2025. For each species, median growth sensitivity +/- standard deviation is reported across the individuals analysed in 2010, 2015, and 2020 - the three drought years considered.

```{r summary table, echo=F, warning=F, message=F}
library(tidyverse)
datasets <- readRDS("../data/HKK-dendro/sensitivity_data.RData")

tree.time <- datasets$tree.time
tree_vars <- datasets$tree_vars
sp_vars <- datasets$sp_vars

sp_new <- read.csv("../data/HKK-dendro/species.csv")

colnames(sp_vars)[1] <- "Species"
sp_vars$iucn <- sp_new$iucn[match(sp_vars$Species, sp_new$sp)]

# merge tree_vars and sp_vars
tree_vars <- merge(tree_vars, sp_vars, by = "Species", all.x = TRUE)
# merge tree_vars and tree.time
tree.time <- merge(tree.time, tree_vars, by = "Tag", all.x = TRUE)

# colnames(tree.time)

# standardise the variables within species and across all--------------------

tree.time <- tree.time %>%
  # remove large outliers for each year
  group_by(yr) %>%
  dplyr::mutate(sens.prop = ifelse(mean(sens.prop, na.rm = TRUE) + 4 * sd(sens.prop, na.rm = TRUE) > sens.prop & sens.prop > mean(sens.prop, na.rm = TRUE) - 4 * sd(sens.prop, na.rm = TRUE), sens.prop, NA)) %>%
  filter(!is.na(sens.prop) & !is.na(cii_min1) & !is.na(calcDBH_min1) & !is.na(twi)) %>%
  ungroup()

# make a species table for each year

sens_sp <- tree.time %>%
  filter(yr %in% c(2010, 2015, 2020)) %>%
  group_by(spfull, yr) %>%
  dplyr::summarise(
    n = n(),
    med_sens = round(median(sens.prop, na.rm = T), 2),
    sd_sens = round(sd(sens.prop, na.rm = T), 2)
  ) %>%
  ungroup()

sens_sp <- merge(sens_sp, dplyr::select(sp_vars, c(spfull, Species, williams_dec, iucn)), by = "spfull", all.x = T)

# head(sens_sp)
# make the column
sens_sp$sens <- paste0(sens_sp$med_sens, " +/- ", sens_sp$sd_sens)

# pivot wider by year

sens_sp <- sens_sp %>%
  dplyr::select(-med_sens, -sd_sens, -n) %>%
  pivot_wider(values_from = sens, names_from = yr) %>%
  dplyr::rename(
    deciduousness = williams_dec,
    "IUCN status" = iucn,
    "Species name" = spfull,
    "Species code" = Species
  )
# head(sens_sp)

knitr::kable(sens_sp)
```
